---
title: "using Decision Trees to predict if a Hotel Reservation will get cancelled"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
author: Sohaib Syed
---


### Part 2.1-A
```{r}
df<-read.csv(file="hotel_bookings.csv")
H1_freq<-sum(with(df,hotel=="Resort Hotel"))
H2_freq<-sum(with(df,hotel=="City Hotel"))
paste("There are", H1_freq, "observations for type H1 hotels")
paste("There are", H2_freq, "observations for type H2 hotels")
```
### Part 2.1-B
```{r}
canceled_freq<-sum(with(df,is_canceled==1))
not_cancelled_freq<-sum(with(df,is_canceled==0))
paste("Number of guests who canceled reservation:", canceled_freq)
paste("Number of guests who did not cancel the reservation:", not_cancelled_freq)
```
### Part 2.1-C
```{r}
max_type<-max(df$customer_type)
num_max_type<-sum(df$customer_type=="Transient-Party")
paste("Customer type with the most reservations is", max_type, "with", num_max_type, "reservations")
```
### Part 2.1-D
```{r}
max_spaces<-max(df$required_car_parking_spaces)
num_max_spaces<-sum(with(df,required_car_parking_spaces==max_spaces))
paste( num_max_spaces, "customers required the most number of parking spaces (",max_spaces,").")
```
### Part 2.1-E
```{r}
min_spaces<-min(df$required_car_parking_spaces)
num_min_spaces<-sum(with(df,required_car_parking_spaces==min_spaces))
paste( num_min_spaces, "customers required the least number of parking spaces (",min_spaces,").")
```
### Part 2.1-F
```{r}
matching_room<-sum(with(df, reserved_room_type==assigned_room_type))
percent_matching<-signif(((matching_room/nrow(df))*100),digits=4)
paste(percent_matching, "of the people who expressed a room preference during reservation got the room during check-in.")
```
### Part 2.1-G
```{r}
city_hotels<- subset(df, hotel=="City Hotel" & country != "NULL"  )
top10_count_forH2<-tail(sort(table(city_hotels$country)),n=10)
barplot(top10_count_forH2,col=rainbow(n=10),ylab = "Number of City Hotels",xlab="Country",main= "10 Countries with most City Hotel Bookings")

resort_hotels<-subset(df,hotel=="Resort Hotel" & country != "NULL")
top10_count_forH1<-tail(sort(table(resort_hotels$country)),n=10)
barplot(top10_count_forH1,col=rainbow(n=10),ylab = "Number of Resort Hotels",xlab="Country",main= "10 Countries with most Resort Hotel Bookings")
```
### Part 2.1-H-i
```{r}
paste("The country with most visitors to either type of hotel is: Portugal")
```
### Part 2.1-H-ii
It can be inferred that the data originates from Portugal. It seems like it might be data that a booking company based in Portugal has which explains why most bookings are from Portugal.

### Part 2.2
```{r}
library("ISLR")
set.seed(1122)
index <- sample(1:nrow(df), .90*dim(df)[1])
train.df <-df[index,]
test.df <-df[-index,]
```
### Part 2.2-A-i
```{r}
library(rpart)
library(rpart.plot)

model<-rpart(is_canceled~lead_time+market_segment+distribution_channel+previous_cancellations+deposit_type+customer_type+required_car_parking_spaces+total_of_special_requests, method="class",data=train.df)

rpart.plot(model, extra=104, fallen.leaves=T, type=4, main="Hotel Booking Decsion Tree")
```
### Part 2.2-A-ii
```{r}
#The most important variables are: deposit_type,total_of_special_requests,lead_time,previous_cancellations,market_segment,customer_type,distribution_channel,required_car_parking_spaces,
```
### Part 2.2-A-iii
```{r}
library(caret)
my_prediction<-predict(model,test.df,type="class")
myMat<-confusionMatrix(my_prediction,as.factor(test.df$is_canceled),positive='1')
myMat
paste("Accuracy:" ,.809, "Error:",.191,"Balanced Accuracy:", .768, "Specificity:",.930,"Sensitivity:",.605,"Precision:",.838)
```
### Part 2.2-A-iv (Code structure courtesy of Dr. Gurbani: diabetes.Rmd)
```{r}
library(ROCR)
pred.rocr<-predict(model,newdata = test.df,type="prob")[,2]
f.pred<-prediction(pred.rocr, test.df$is_canceled)
f.perf <- performance(f.pred, "tpr", "fpr")
plot(f.perf, colorize=T, lwd=3)
abline(0,1)
```
### Part 2.2-A-v  (Code structure courtesy of Dr. Gurbani: diabetes.Rmd)
```{r}
auc <- performance(f.pred, measure = "auc")
cat(paste("The area under curve (AUC) for this model is ", round(auc@y.values[[1]], 3)))
```
### Part 2.3-A
```{r}
set.seed(1122)
index <- sample(1:nrow(df), .90*dim(df)[1])
new_train.df <-df[index,]
new_test.df <-df[-index,]

new_model<-rpart(is_canceled~lead_time+market_segment+distribution_channel+previous_cancellations+deposit_type+customer_type+required_car_parking_spaces+total_of_special_requests,method="class",data=new_train.df,control = rpart.control(cp = 0.0))

new_pred<-predict(new_model,new_test.df,type="class")

new_Mat<-confusionMatrix(new_pred,as.factor(new_test.df$is_canceled),positive='1')

new_Mat
cat("Before Pruning:\n  Accuracy: 0.821\n  Error: 0.179\n  Balanced Acc.: 0.793\n  Specificity: 0.904\n  Sensitivity: 0.682\n  Precision: 0.808")
```
### Part 2.3-B
```{r}
cpx=new_model$cptable[which.min(new_model$cptable[,"xerror"]), "CP"] #(Code courtesy of Dr. Gurbani: diabetes.Rmd)
ptree<-prune(new_model,cp=cpx)
paste("Prune point occurs at a complexity of", round(cpx,5))
paste("At this complexity, xerror is",round(min(new_model$cptable[,"xerror"]),5))
```
### Part 2.3-C
```{r}
pruned_pred<-predict(ptree,new_test.df,type="class")
prune_Mat<-confusionMatrix(pruned_pred,as.factor(new_test.df$is_canceled),positive='1')
prune_Mat
cat("After Pruning:\n  Accuracy: 0.823\n  Error: 0.177\n  Balanced Acc.: 0.792\n  Specificity: 0.916\n  Sensitivity: 0.668\n  Precision: 0.825")
```
### Part 2.3-D
```{r}
cat("The pruned tree generalzies better.")
```
### Part 2.3-E
```{r}
cat("The best tree is the pruned tree because it has the least error and is still almost as precise as the tree in 2.2")
```
